---
title: "Decennial Race Data Pull"
author: "Anthony Gallegos"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
---

```{r knitr_init, echo=FALSE, cache=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(pander)
library(devtools)
library(censusapi)
library(maptools)
library(rgdal)
library(geojsonio)
library(rmapshaper)
library(sf)
library(ggplot2)
library(tidyr)
library(viridis)
## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```


# Data Pull Code for Race

## 1990,2000,2010 Decennials using race as an example. 1990 has the most limiting data collected regarding race,
## and the 2000 and 2010 data is transformed to reflect it (the totals equal to the total, suggesting the data
## transformation does not adversely impact the integrity of the data).


```{r}

# Set Working Directory
setwd("~/Documents/Capstone/Solo Project/Markdown")

# Census Key
censuskey <- "e2a75a26c7d4adf9db476fa739110c582113d15e"

```

## 1990 Data Pull

```{r}

# Read csv file with only race variables. I manually created this by cutting and pasting the rows I wanted from the previous spreadsheet. 

`1990_race` <- read.csv("/home/anthony/Documents/Capstone/Solo Project/Variables/race_1990.csv")

# Create Variable list based off variables in the first column
var.list.race.1990 <- `1990_race`$name


# Pull Data using variable list only. Make sure the censusapi package is loaded.
race.pop.1990 <- getCensus( name = "sf3", vintage = 1990, key=censuskey,
                               vars = var.list.race.1990,
                               region="tract:*", regionin="state:04&county:13" )

# Rename data
race.pop.1990 <- rename(race.pop.1990, Total_1990=P0010001, White_1990=P0080001, Black_1990=P0080002, Indian_1990=P0080003, Asian_or_Pacific_Islander_1990=P0080004, Other_Race_1990=P0080005)

# I will need to use the paste function to identify the year for these variables.

# Write into new csv file
write.csv( race.pop.1990, "/home/anthony/Documents/Capstone/Solo Project/Written CSVs/race_1990_pop.csv", row.names=T)




```


## 2000 Data Pull
```{r}


# Read csv file with only race variables. I manually created this by cutting and pasting the rows I wanted from the previous spreadsheet. 

`2000_race` <- read.csv("/home/anthony/Documents/Capstone/Solo Project/Variables/race_2000.csv")

# Create Variable list based off variables in the first column
var.list.race.2000 <- `2000_race`$name


# Pull Data using variable list only. Make sure the censusapi package is loaded.
race.pop.2000 <- getCensus( name = "sf3", vintage = 2000, key=censuskey,
                               vars = var.list.race.2000,
                               region="tract:*", regionin="state:04&county:13" )

# Rename data
race.pop.2000 <- rename(race.pop.2000, Total_2000=P006001, White_2000=P006002, Black_2000=P006003, Indian_2000=P006004, Asian_2000=P006005, Pacific_Islander_2000=P006006, Other_Race_Single_2000=P006007, Other_Race_Multiple_2000=P006008 )

# Group by 
race.pop.2000.mod <- race.pop.2000 %>% group_by(tract)


# Sum Other_Race_Single and Other_Race_Multiple and create new column

race.pop.2000.mod <-race.pop.2000.mod %>% mutate( Other_Race_2000 = Other_Race_Single_2000 + Other_Race_Multiple_2000)


# Sum Asian and Pacific_Islander and create new column

race.pop.2000.mod <-race.pop.2000.mod %>% mutate( Asian_or_Pacific_Islander_2000 = Asian_2000 + Pacific_Islander_2000)


# Remove Other_Race_Single, Other_Race_Multiple, Asian, and Pacific_Islander columns

race.pop.2000.mod <- subset(race.pop.2000.mod, select = -c(Asian_2000, Pacific_Islander_2000, Other_Race_Single_2000, Other_Race_Multiple_2000 ))

# I will need to use the paste function to identify the year for these variables.

# Write into new csv file
write.csv( race.pop.2000.mod, "/home/anthony/Documents/Capstone/Solo Project/Written CSVs/race_2000_pop.csv", row.names=T)



```

## 2010 Data Pull
```{r}

# Read csv file with only race variables. I manually created this by cutting and pasting the rows I wanted from the previous spreadsheet. 

`2010_race` <- read.csv("/home/anthony/Documents/Capstone/Solo Project/Variables/race_2010.csv")

# Create Variable list based off variables in the first column
var.list.race.2010 <- `2010_race`$name


# Pull Data using variable list only. Make sure the censusapi package is loaded.
race.pop.2010 <- getCensus( name = "dec/sf1", vintage = 2010, key=censuskey,
                               vars = var.list.race.2010,
                               region="tract:*", regionin="state:04&county:013" )


# Rename data
race.pop.2010 <- rename(race.pop.2010, Total_2010=H008001, White_2010=H008002, Black_2010=H008003, Indian_2010=H008004, Asian_2010=H008005, Pacific_Islander_2010=H008006, Other_Race_2010=H008007 )

# Group by 
race.pop.2010.mod <- race.pop.2010 %>% group_by(tract)

# Sum Asian and Pacific_Islander and create new column

race.pop.2010.mod <-race.pop.2010.mod %>% mutate( Asian_or_Pacific_Islander_2010 = Asian_2010 + Pacific_Islander_2010)


# Remove Other_Race_Single, Other_Race_Multiple, Asian, and Pacific_Islander columns

race.pop.2010.mod <- subset(race.pop.2010.mod, select = -c(Asian_2010, Pacific_Islander_2010 ))

# I will need to use the paste function to identify the year for these variables.

# Write into new csv file
write.csv( race.pop.2010.mod, "/home/anthony/Documents/Capstone/Solo Project/Written CSVs/race_2010_pop.csv", row.names=T)



```

## Import Shapefile and Join to Data

```{r}
# The commands below work, but this is only for the 2010 data. Need to take new dataset that is combined for all 3 decennials
# instead of joining with the 2010 data.

county_shp <- "//home/anthony/Documents/Capstone/Solo Project/Shape Files/tl_2010_04013_tract10.shp"

# Read shape file and create and create variable for it
county_st <- st_read(county_shp)

# Join shapefile with race by tract
mcpa <- left_join(county_st, race.pop.2010.mod, by=c("TRACTCE10"="tract"))


```

## Percentages (Not complete)

```{r}

# Add Percentages (probably makes more sense to do this after combining everything)

# race.pop <- race.pop %>% group_by(tract) %>% mutate(( White_Percent = White / Total) * 100) %>%
# mutate(( Black_Percent = Black / Total) * 100) %>% 
# mutate(( Indian_Percent = American_Indian_Eskimo_or_Aleu / Total) * 100) %>%
# mutate(( Asian_Percent = Asian_or_Pacific_Islander / Total) * 100) %>%
# mutate(( Other_Percent = Other_Race / Total) * 100)




```

