---
title: "Education Level"
author: "Anthony Gallegos"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    theme: flatly
    highlight: haddock
    toc: yes
    toc_float: yes
---

```{r knitr_init, echo=FALSE, cache=FALSE, message=FALSE}
library(knitr)
library(dplyr)
library(pander)
library(devtools)
library(censusapi)
library(maptools)
library(rgdal)
library(geojsonio)
library(rmapshaper)
library(sf)
library(ggplot2)
library(tidyr)
library(viridis)
## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```


# Data Pull Code for Education Level

This is going to have to be done at the lowest common denometer. In this case, the 1990 data is the least robust.

I'm going to need to do the following to collapse the data into like categories.

2000 Data. Same categories, but I will need to pull both male and female versions and combine them. Simple Enough.

2010 Data. Even more robust categories. I'm going to need to collapse the variables even more and then combine male/female.




## This is not using the 2010 sf1 data, as it is not included. Instead, the acs/ac5 data will be used.

```{r}

# Set Working Directory
setwd("~/Documents/Capstone/Solo Project/Markdown")

# Census Key
censuskey <- "e2a75a26c7d4adf9db476fa739110c582113d15e"

```

## 1990 Education Level Data Pull
```{r}

# Var Pull

`1990_edu_lvl` <- read.csv("/home/anthony/Documents/Capstone/Solo Project/Variables/edu_lvl/edu_1990.csv")

# Create Variable list based off variables in the first column
var.edu.1990 <- `1990_edu_lvl`$name


# Pull Data using variable list only. Make sure the censusapi package is loaded.
edu.1990 <- getCensus( name = "sf3", vintage = 1990, key=censuskey,
                               vars = var.edu.1990,
                               region="tract:*", regionin="state:04&county:013" )



# Rename data
edu.lvl.1990 <- rename( edu.1990, Less_9th_1990=P0570001, More_9th_1990=P0570002, High_Sch_Grad_1990=P0570003, Some_College_1990=P0570004, Assc_Deg_1990=P0570005, Bachelors_Deg_1990=P0570006, Grad_Deg_1990=P0570007 )
```


## 2000 Education Level Data Pull
```{r}

# Var Pull

`2000_edu_lvl` <- read.csv("/home/anthony/Documents/Capstone/Solo Project/Variables/edu_lvl/edu_2000.csv")

# Create Variable list based off variables in the first column
var.edu.2000 <- `2000_edu_lvl`$name


# Pull Data using variable list only. Make sure the censusapi package is loaded.
edu.2000 <- getCensus( name = "sf3", vintage = 2000, key=censuskey,
                               vars = var.edu.2000,
                               region="tract:*", regionin="state:04&county:013" )


# New var combining male and female

edu.2000$Less_9th_2000 <- (edu.2000$P148A003 + edu.2000$P148A011 )

edu.2000$More_9th_2000 <- (edu.2000$P148A004 + edu.2000$P148A012  )

edu.2000$High_Sch_Grad_2000 <- (edu.2000$P148A005 + edu.2000$P148A013)

edu.2000$Some_College_2000 <- (edu.2000$P148A006 + edu.2000$P148A014)

edu.2000$Assc_Deg_2000 <- (edu.2000$P148A007 + edu.2000$P148A015 )

edu.2000$Bachelors_Deg_2000 <- (edu.2000$P148A008 + edu.2000$P148A016)

edu.2000$Grad_Deg_2000 <- (edu.2000$P148A008 + edu.2000$P148A017 )


```


## 2010 ACS Data Pull
```{r}

# It may be necessary to pull the 2010 ACS because the 2010 decennial sf1 data is not nearly as comprehensive.

`2010_edu_lvl` <- read.csv("/home/anthony/Documents/Capstone/Solo Project/Variables/edu_lvl/edu_2010.csv")

# Create Variable list based off variables in the first column
var.edu.2010 <- `2010_edu_lvl`$name
 

# Pull Data using variable list only. Make sure the censusapi package is loaded.
edu.2010acs <- getCensus( name = "acs/acs5", vintage = 2010, key=censuskey,
                               vars = var.edu.2010,
                               region="tract:*", regionin="state:04&county:013" )



# New var combining male and female

edu.2010acs$Less_9th_2010 <- (edu.2010acs$B15002_003E + edu.2010acs$B15002_004E + edu.2010acs$B15002_005E + edu.2010acs$B15002_006E +  edu.2010acs$B15002_020E + edu.2010acs$B15002_021E + edu.2010acs$B15002_022E + edu.2010acs$B15002_023E   )

edu.2010acs$More_9th_2010 <- (edu.2010acs$B15002_007E + edu.2010acs$B15002_008E + edu.2010acs$B15002_009E + edu.2010acs$B15002_010E + edu.2010acs$B15002_025E + edu.2010acs$B15002_025E + edu.2010acs$B15002_026E + edu.2010acs$B15002_027E + edu.2010acs$B15002_024E )

edu.2010acs$High_Sch_Grad_2010 <- (edu.2010acs$B15002_011E + edu.2010acs$B15002_028E )

edu.2010acs$Some_College_2010 <- (edu.2010acs$B15002_012E + edu.2010acs$B15002_013E + edu.2010acs$B15002_029E + edu.2010acs$B15002_030E)

edu.2010acs$Assc_Deg_2010 <- (edu.2010acs$B15002_014E + edu.2010acs$B15002_031E )

edu.2010acs$Bachelors_Deg_2010 <- (edu.2010acs$B15002_015E + edu.2010acs$B15002_032E )

edu.2010acs$Grad_Deg_2010 <- (edu.2010acs$B15002_016E + edu.2010acs$B15002_017E + edu.2010acs$B15002_018E + edu.2010acs$B15002_033E + edu.2010acs$B15002_034E + edu.2010acs$B15002_035E)











```





